---
title: "Mapping domains to the human-viral protein interaction network"
author: "Vitalii Kleshchevnikov"
date: "29/06/2017"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages = c("data.table","downloader","R.utils","UniProt.ws", "PSICQUIC", "ggplot2", "rtracklayer", "Biostrings", "MItools", "GGally")
if(mean(packages %in% names(installed.packages()[,"Package"])) != 1){
    packages_to_install = packages[!packages %in% names(installed.packages()[,"Package"])]
    # specifying mirror is necessary for some Linux systems
    install.packages(packages_to_install, dependencies = T, repos = "http://mirrors.ebi.ac.uk/CRAN/")
    packages_to_install = packages[!packages %in% names(installed.packages()[,"Package"])]
    source("https://bioconductor.org/biocLite.R")
    biocLite(packages_to_install)
    devtools::install_github("vitkl/MItools")
}

suppressPackageStartupMessages({
    library(data.table)
    library(downloader)
    library(R.utils)
    library(UniProt.ws)
    library(PSICQUIC)
    library(ggplot2)
    library(rtracklayer)
    library(Biostrings)
    library(MItools)
    library(GGally)
})
```

Date: `r Sys.time()`

## Map domain information on the protein interaction network

I read interaction data and clean this data to make it more useble. Then, I filter and keep only human-viral interactions. All done by the interSpeciesInteractome function.  

```{r human_viral}
# load protein-domain mapping
protein_domain_pair = fread(file = "./processed_data_files/protein_domain_pair", stringsAsFactors = F, sep = "\t")
# load PPI data
all_human_interaction = fullInteractome(taxid = 9606, database = "IntActFTP", format = "tab27",
                                        clean = TRUE, protein_only = TRUE, directory = "./data_files/")
all_viral_interaction = interSpeciesInteractome(taxid1 = 9606, taxid2 = 10239, database = "IntActFTP", format = "tab27",
                                        clean = TRUE, protein_only = TRUE, directory = "./data_files/")
all_within_viral_interaction = fullInteractome(taxid = 10239, database = "IntActFTP", format = "tab27",
                                        clean = TRUE, protein_only = TRUE, directory = "./data_files/")
all_within_viral_interaction = subsetMITABbyID(MITABdata = all_within_viral_interaction, ID_seed = unique(all_viral_interaction$data$IDs_interactor_B), within_seed = T, only_seed2nonseed = F)
BioPlex3 = loadBioplex(dir = "./data_files/",
                       url = "http://bioplex.hms.harvard.edu/data/BioPlex_2.3_interactionList.tsv")

# N domains
protein_domain_pair[IDs_protein %in% unique(all_viral_interaction$data$IDs_interactor_A), uniqueN(IDs_domain)]
# Pfam domains 
protein_domain_pair[IDs_protein %in% unique(all_viral_interaction$data$IDs_interactor_A) & grepl("PF",all_IDs_domain), uniqueN(IDs_domain)]
qplot(protein_domain_pair[IDs_protein %in% unique(all_viral_interaction$data$IDs_interactor_A)][, .(N = uniqueN(IDs_domain)), by = IDs_protein]$N, main = "number of domains per protein", xlab = "number of domains per protein", geom = "histogram") + scale_x_log10() + theme_light()

qplot(protein_domain_pair[IDs_protein %in% unique(all_viral_interaction$data$IDs_interactor_A)][, .(N = .N), by = IDs_domain]$N, main = "domain occurence", xlab = "number of proteins per domain", geom = "histogram") + scale_x_log10() + theme_light()
```

Viral proteins have `r length(unique(all_viral_interaction$data$pair_id))` known interacting pairs with human proteins while only `r length(unique(all_within_viral_interaction$data$pair_id))` with other viral proteins.  

Both the interaction network table and the domain data contain more information than necessary for identifying domains likely to mediate interaction. Selecting only what's necessary: viral_protein_UniprotID - human_protein_UniprotID - human_domain_InterProID.  

```{r simplify_interactions}
all_viral_interaction_simp = unique(all_viral_interaction$data[,.(IDs_interactor_human = IDs_interactor_A,
                                                                  IDs_interactor_viral = IDs_interactor_B,
                                                                  Taxid_interactor_human = Taxid_interactor_A,
                                                                  Taxid_interactor_viral = Taxid_interactor_B)])
# human interactions need simplyfying
all_human_interaction_simp = unique(all_human_interaction$data[,.(Taxid_interactor_human_A = Taxid_interactor_A,
                                                                  Taxid_interactor_human_B = Taxid_interactor_B,
                                                                  pair_id = pair_id)])
# but also expanding into directional network (both A -> B and B -> A)
all_human_interaction_simp[, c("IDs_interactor_human_A", "IDs_interactor_human_B") := tstrsplit(pair_id, "\\|")]
all_human_interaction_simp = rbind(all_human_interaction_simp,
                                   all_human_interaction_simp[,.(IDs_interactor_human_A = IDs_interactor_human_B,
                                                                 IDs_interactor_human_B = IDs_interactor_human_A,
                                                                 Taxid_interactor_human_A = Taxid_interactor_human_B,
                                                                  Taxid_interactor_human_B = Taxid_interactor_human_A,
                                                                  pair_id = pair_id)])
all_human_interaction_simp[, pair_id := NULL]
all_human_interaction_simp = unique(all_human_interaction_simp)

# Human interactions from BioPlex3 BioPlex3
BioPlex3_human_interaction_simp = unique(BioPlex3$data[,.(Taxid_interactor_human_A = Taxid_interactor_A,
                                                                  Taxid_interactor_human_B = Taxid_interactor_B,
                                                                  pair_id = pair_id)])
# but also expanding into directional network (both A -> B and B -> A)
BioPlex3_human_interaction_simp[, c("IDs_interactor_human_A", "IDs_interactor_human_B") := tstrsplit(pair_id, "\\|")]
BioPlex3_human_interaction_simp = rbind(BioPlex3_human_interaction_simp,
                                   BioPlex3_human_interaction_simp[,.(IDs_interactor_human_A = IDs_interactor_human_B,
                                                                 IDs_interactor_human_B = IDs_interactor_human_A,
                                                                 Taxid_interactor_human_A = Taxid_interactor_human_B,
                                                                  Taxid_interactor_human_B = Taxid_interactor_human_A,
                                                                  pair_id = pair_id)])
BioPlex3_human_interaction_simp[, pair_id := NULL]
BioPlex3_human_interaction_simp = unique(BioPlex3_human_interaction_simp)
```

Next, the frequency of each domain is calculated and simplified domain information (IDs_protein, IDs_domain, domain_type) is integrated with the simplified network data described above.  

```{r interactions_plus_domains_viral}
protein_domain_pair_temp = copy(protein_domain_pair)[, IDs_interactor_human := IDs_protein][, IDs_protein := NULL][, IDs_domain_human := IDs_domain][, IDs_domain := NULL]

# calculate total number of human proteins
protein_domain_pair_temp[, N_prot_w_interactors := length(unique(IDs_interactor_human))]
# calculate domain count and frequency
protein_domain_pair_temp[, domain_count := length(unique(IDs_interactor_human)), by = IDs_domain_human]
protein_domain_pair_temp[, domain_frequency := domain_count / N_prot_w_interactors]

# calculate network descriptive stats
# viral protein degree
all_viral_interaction_simp[, IDs_interactor_viral_degree := length(unique(IDs_interactor_human)), by = IDs_interactor_viral]
# human protein degree
all_viral_interaction_simp[, IDs_interactor_human_degree := length(unique(IDs_interactor_viral)), by = IDs_interactor_human]

# I keep all interactions even if a human protein has no known domain (all.x = F, all.y = T)
viral_human_w_domains = merge(protein_domain_pair_temp, all_viral_interaction_simp, all.x = F, all.y = T, by = "IDs_interactor_human", allow.cartesian = T)

# human domains per viral protein
viral_human_w_domains[, IDs_domain_human_per_IDs_interactor_viral := length(unique(IDs_domain_human)), by = IDs_interactor_viral]
# viral protein per human domain
viral_human_w_domains[, IDs_interactor_viral_per_IDs_domain_human := length(unique(IDs_interactor_viral)), by = IDs_domain_human]

# domain count but per viral protein human domain instances (how many proteins the domain is located in) per viral protein (ID) and human domain (ID)
viral_human_w_domains[, domain_count_per_IDs_interactor_viral := length(unique(IDs_interactor_human)), by = .(IDs_interactor_viral, IDs_domain_human)]
viral_human_w_domains[is.na(IDs_domain_human), domain_count_per_IDs_interactor_viral := 0]
# domain frequency but per viral protein
viral_human_w_domains[, domain_frequency_per_IDs_interactor_viral := domain_count_per_IDs_interactor_viral / IDs_interactor_viral_degree, by = IDs_interactor_viral]
viral_human_w_domains[is.na(IDs_domain_human), domain_frequency_per_IDs_interactor_viral := 0]
# fold enrichment
viral_human_w_domains[, fold_enrichment := domain_frequency_per_IDs_interactor_viral / domain_frequency]
viral_human_w_domains[is.na(IDs_domain_human), fold_enrichment := 0]


# save resulting network
fwrite(viral_human_w_domains, file = "./processed_data_files/viral_human_net_w_domains", sep = "\t")
```

```{r interactions_plus_domains_human}
protein_domain_pair_temp = copy(protein_domain_pair)[, IDs_interactor_human_B := IDs_protein][, IDs_protein := NULL][, IDs_domain_human_B := IDs_domain][, IDs_domain := NULL]

# calculate total number of human proteins
protein_domain_pair_temp[, N_prot_w_interactors := length(unique(IDs_interactor_human_B))]
# calculate domain count and frequency
protein_domain_pair_temp[, domain_count := length(unique(IDs_interactor_human_B)), by = IDs_domain_human_B]
protein_domain_pair_temp[, domain_frequency := domain_count / N_prot_w_interactors]

# calculate network descriptive stats
# viral protein degree
all_human_interaction_simp[, IDs_interactor_human_A_degree := length(unique(IDs_interactor_human_B)), by = IDs_interactor_human_A]
# human protein degree
all_human_interaction_simp[, IDs_interactor_human_B_degree := length(unique(IDs_interactor_human_A)), by = IDs_interactor_human_B]

# I keep all interactions even if a human protein has no known domain (all.x = F, all.y = T)
human_w_domains = merge(protein_domain_pair_temp, all_human_interaction_simp, all.x = F, all.y = T, by = "IDs_interactor_human_B", allow.cartesian = T)

# human domains per viral protein
human_w_domains[, IDs_domain_human_B_per_IDs_interactor_human_A := length(unique(IDs_domain_human_B)), by = IDs_interactor_human_A]
# viral protein per human domain
human_w_domains[, IDs_interactor_human_A_per_IDs_domain_human_B := length(unique(IDs_interactor_human_A)), by = IDs_domain_human_B]

# domain count but per viral protein human domain instances (how many proteins the domain is located in) per viral protein (ID) and human domain (ID)
human_w_domains[, domain_count_per_IDs_interactor_human_A := length(unique(IDs_interactor_human_B)), by = .(IDs_interactor_human_A, IDs_domain_human_B)]
human_w_domains[is.na(IDs_domain_human_B), domain_count_per_IDs_interactor_viral := 0]
# domain frequency but per viral protein
human_w_domains[, domain_frequency_per_IDs_interactor_human_A := domain_count_per_IDs_interactor_human_A / IDs_interactor_human_A_degree, by = IDs_interactor_human_A]
human_w_domains[is.na(IDs_domain_human_B), domain_frequency_per_IDs_interactor_human_A := 0]
# fold enrichment
human_w_domains[, fold_enrichment := domain_frequency_per_IDs_interactor_human_A / domain_frequency]
human_w_domains[is.na(IDs_domain_human_B), fold_enrichment := 0]


# save resulting network
fwrite(human_w_domains, file = "./processed_data_files/human_net_w_domains", sep = "\t")
gzip("./processed_data_files/human_net_w_domains", remove = T, overwrite = T)
```

```{r interactions_plus_domains_human_BioPlex3}
# calculate network descriptive stats
# viral protein degree
BioPlex3_human_interaction_simp[, IDs_interactor_human_A_degree := length(unique(IDs_interactor_human_B)), by = IDs_interactor_human_A]
# human protein degree
BioPlex3_human_interaction_simp[, IDs_interactor_human_B_degree := length(unique(IDs_interactor_human_A)), by = IDs_interactor_human_B]

# I keep all interactions even if a human protein has no known domain (all.x = F, all.y = T)
BioPlex3human_w_domains = merge(protein_domain_pair_temp, BioPlex3_human_interaction_simp, all.x = F, all.y = T, by = "IDs_interactor_human_B", allow.cartesian = T)

# human domains per viral protein
BioPlex3human_w_domains[, IDs_domain_human_B_per_IDs_interactor_human_A := length(unique(IDs_domain_human_B)), by = IDs_interactor_human_A]
# viral protein per human domain
BioPlex3human_w_domains[, IDs_interactor_human_A_per_IDs_domain_human_B := length(unique(IDs_interactor_human_A)), by = IDs_domain_human_B]

# domain count but per viral protein human domain instances (how many proteins the domain is located in) per viral protein (ID) and human domain (ID)
BioPlex3human_w_domains[, domain_count_per_IDs_interactor_human_A := length(unique(IDs_interactor_human_B)), by = .(IDs_interactor_human_A, IDs_domain_human_B)]
BioPlex3human_w_domains[is.na(IDs_domain_human_B), domain_count_per_IDs_interactor_viral := 0]
# domain frequency but per viral protein
BioPlex3human_w_domains[, domain_frequency_per_IDs_interactor_human_A := domain_count_per_IDs_interactor_human_A / IDs_interactor_human_A_degree, by = IDs_interactor_human_A]
BioPlex3human_w_domains[is.na(IDs_domain_human_B), domain_frequency_per_IDs_interactor_human_A := 0]
# fold enrichment
BioPlex3human_w_domains[, fold_enrichment := domain_frequency_per_IDs_interactor_human_A / domain_frequency]
BioPlex3human_w_domains[is.na(IDs_domain_human_B), fold_enrichment := 0]


# save resulting network
fwrite(BioPlex3human_w_domains, file = "./processed_data_files/BioPlex3human_net_w_domains", sep = "\t")
gzip("./processed_data_files/BioPlex3human_net_w_domains", remove = T, overwrite = T)
```

## Summary of the viral-human network

The plot below show the relationships (2D histogram), distribution density (on the diagonal) and pearson correlation for a number of parameters characterising human domains, viral proteins or it's interactions:

1. domain_frequency is the number of human proteins with a particular domain divided by the total number of human proteins (the attribute of a human domain)
2. IDs_interactor_viral_degree is the number of interactions each viral protein has (the attribute of a viral protein)
3. IDs_interactor_human_degree is the number of interactions each human protein has (the attribute of a human protein)
4. IDs_domain_human_per_IDs_interactor_viral is the number of domain types that are present in proteins which a particular viral protein interacts with (the attribute of a viral protein)

```{r, fig.height=9, fig.width=13}
# function to accomodate ggplot2::geom_bin2d in GGally::ggpairs, taken from http://ggobi.github.io/ggally/#custom_functions
d2_bin <- function(data, mapping, ..., low = "#132B43", high = "#56B1F7") {
    ggplot(data = data, mapping = mapping) +
        geom_bin2d(...) +
        scale_fill_gradient(low = low, high = high) +
        scale_y_log10() + scale_x_log10() + annotation_logticks()
}

log10_density = function(data, mapping, ...){
    ggplot(data = data, mapping = mapping) +
        geom_density(...) +
        scale_x_log10() + annotation_logticks()
}

d2_bin_plot = GGally::ggpairs(viral_human_w_domains[!is.na(IDs_domain_human),.(domain_frequency, 
                                                                 IDs_interactor_viral_degree, 
                                                                 IDs_interactor_human_degree, 
                                                                 IDs_domain_human_per_IDs_interactor_viral, 
                                                                 IDs_interactor_viral_per_IDs_domain_human,
                                                                 domain_count_per_IDs_interactor_viral,
                                                                 domain_frequency_per_IDs_interactor_viral,
                                                                 fold_enrichment)], 
                lower = list(continuous = d2_bin), 
                diag = list(continuous = log10_density)) +
    theme_light() +
    theme(strip.text.y = element_text(angle = 0, size = 10),
          strip.text.x = element_text(angle = 90, size = 10))
d2_bin_plot
```

## Summary of the viral-human network

## R session information

```{r}
save(list = ls(), file="./processed_data_files/map_domains_to_human_viral_network_clust.RData")
R.utils::gzip(filename = "./processed_data_files/map_domains_to_human_viral_network_clust.RData",
              destname = "./processed_data_files/map_domains_to_human_viral_network_clust.RData.gz",
              remove = T, overwrite = T)

Sys.Date()
sessionInfo()
```