---
title: "Motif_search_strategies"
author: "Vitalii Kleshchevnikov"
date: "29/11/2017"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages = c("MItools", "rtracklayer", "R.utils", "ontologyIndex", "parallel")
if(mean(packages %in% names(installed.packages()[,"Package"])) != 1){
    packages_to_install = packages[!packages %in% names(installed.packages()[,"Package"])]
    # specifying mirror is necessary for some Linux systems
    install.packages(packages_to_install, dependencies = T, repos = "http://mirrors.ebi.ac.uk/CRAN/")
    packages_to_install = packages[!packages %in% names(installed.packages()[,"Package"])]
    source("https://bioconductor.org/biocLite.R")
    biocLite(packages_to_install)
    devtools::install_github("vitkl/MItools", dependencies = T)
}

suppressPackageStartupMessages({
    library(MItools)
    library(rtracklayer)
    library(parallel)
})
```

# Introduction

## Strategies

1. Optimal search settings (3*2):  
- slimlen= 5,8,10 : Maximum length of SLiMs to return (no. non-wildcard positions)  
- maxwild= 2,3 : Maximum number of consecutive wildcard positions to allow  

2. Comparing interaction datasets (4):  
- Full IntAct  
- BioPlex  
- Vidal  
- randomised BioPlex (keeping the degree and the number of interactions intact)  

3. Domain-centered vs protein-centered strategy (2)  

4. QSLIMFinder vs SLIMFinder (2)  

Total: 3 \* 2 \* 4 \* 2 \* 2 = 96 conditions  

This is not feasible, so I will first compare interaction datasets, domain- vs protein- centered strategy and QSLIMFinder vs SLIMFinder (16 conditions).  
I will then test multiple motif settings on a selected dataset (different run of the pipeline, 6 conditions).  

## load PPI data

```{r load_data}
# load ppi data
IntAct = loadIntActFTP(dir = "./data_files/IntActRelease_2017Nov13/")
# human-viral
all_viral_interaction = interSpeciesInteractome(taxid1 = 9606, taxid2 = 10239, database = "IntActFTP", format = "tab27",
                                                clean = TRUE, protein_only = TRUE,
                                                MITABdata = IntAct, directory = "./data_files/")
# human-human
Full_IntAct = fullInteractome(taxid = 9606, database = "IntActFTP", format = "tab27",
                              clean = TRUE, protein_only = TRUE,
                              MITABdata = IntAct, directory = "./data_files/")
BioPlex = loadBioplex(dir = "./data_files/",
                      url = "http://bioplex.hms.harvard.edu/data/BioPlex_2.3_interactionList.tsv")
Vidal = subsetMITABbyPMIDs(MITABdata = Full_IntAct,
                           PMIDs = c("25416956", "unassigned1304"))

randomised_BioPlex = copy(BioPlex)
randomised_BioPlex$data$Publication_Identifiers = NA
randomised_BioPlex$data$Confidence_values = NA
randomised_BioPlex$data = unique(randomised_BioPlex$data)
set.seed(1)
randomised_BioPlex$data$IDs_interactor_B = randomised_BioPlex$data$IDs_interactor_B[sample.int(length(randomised_BioPlex$data$IDs_interactor_B))]
randomised_BioPlex$data$pair_id = paste0(randomised_BioPlex$data$IDs_interactor_A, "|", randomised_BioPlex$data$IDs_interactor_B)
```

## Set standard options

```{r std_options}
seed_list = c("O00264","O00264-PRO_0000038428","O00469","O14545","O14979","O15031","O15269","O15270","O43149","O43164","O43294","O60518","O60825","O75477","O95639","P00403","P01892","P02795","P05023","P05546-PRO_0000037940","P05546-PRO_0000037946","P06576","P06703","P08195","P08962","P09543","P11166","P13667","P16615","P16615-PRO_0000045596","P16989","P21912","P21964","P27824","P35613","P35658","P38435","P49257","P49755","P50395","P50395-PRO_0000037966","P51570","P51648-PRO_0000045596","P51648","P51659","P52630","P52756","P54136","P61201","P61619","P63151","P82979","Q00587","Q02539","Q03188-PRO_0000037551","Q03518","Q08999","Q13105","Q13586","Q13724","Q13724-PRO_0000045596","Q14677","Q14789","Q15369","Q49A26","Q53H12","Q5C9Z4","Q5H9R7","Q5HYI8","Q5SXM2","Q5T3F8-PRO_0000037548","Q5VTE0","Q86YA3","Q8IZY2","Q8N448","Q8N7W2","Q8NBS9-PRO_0000045602","Q8NI60","Q8TED0","Q8WVX9","Q92538","Q92805","Q969X5","Q96ST2","Q96T23","Q96TA2","Q99442","Q9BQ70","Q9BX40","Q9H6S0-PRO_0000037576","Q9H9L4-PRO_0000037566","Q9NP80-PRO_0000037548","Q9NSY0","Q9NXH8","Q9NZ01","Q9NZM1","Q9UJZ1","Q9UPN7","Q9UPU5","Q9Y3T9-PRO_0000037713","Q9Y4F1","Q9Y4W2","Q9Y5M8","Q9Y6N5")

# set standard options
myPPInetwork2SLIMFinder = function(dataset_name = "SLIMFinder", slimlen = 5, maxwild = 2,
                                   interaction_main_set = Full_IntAct,
                                   center_domains = F, analysis_type = "qslimfinder"){
    PPInetwork2SLIMFinder(dataset_name = dataset_name,
                          interaction_main_set = interaction_main_set,
                          interaction_query_set = all_viral_interaction,
                          analysis_type = analysis_type,
                          options = paste0("dismask=T consmask=F cloudfix=T probcut=0.3 minwild=0 maxwild=",maxwild," slimlen=",slimlen," alphahelix=F maxseq=1500 savespace=1 iuchdir=T"),
                          path2domain_enrich = "./processed_data_files/what_we_find_VS_ELM_clust20171019.RData",
                          domain_enrich_object = "res_count",
                          center_domains = center_domains,
                          fasta_path = "./data_files/all_human_viral_proteins.fasta",
                          main_set_only = F,
                          domain_pvalue_cutoff = 0.000005, # 0.000005 for test run, 1 for full run
                          SLIMFinder_dir = paste0("./",dataset_name,"/"),
                          LSF_project_path = "/hps/nobackup/research/petsalaki/users/vitalii/vitalii/viral_project/",
                          software_path = "../software/cluster/",
                          length_set1_min = 4, # selected based on benchmarking: minimal set size that returns motifs with Sig < 0.05
                          length_set2_min = 1,
                          write_log = T,
                          N_seq = 200,
                          seed_list = seed_list, # NULL for full run
                          memory_start = 350,
                          memory_step = 100)
}
```

## Define parameters

```{r params}
analysis_type = c("qslimfinder", "slimfinder")
datasets = c("Full_IntAct", "BioPlex", "Vidal", "randomised_BioPlex")
domain_centered = c(F, T)
#slimlen = c(5, 8, 10)
#maxwild = c(2, 3)

dataset_names = data.table(analysis_type = character(), datasets = character(),
                           domain_centered = logical(), dataset_names = character(), 
                           slimlen = integer(), maxwild = integer())
for(a in analysis_type){
    for(d in datasets){
        for(c in domain_centered){
            dataset_names_temp = data.table(analysis_type = a, datasets = d,
                                            domain_centered = c, dataset_names = paste(a,d,c, sep = "."), 
                                            slimlen = 5, maxwild = 2)
            dataset_names = rbind(dataset_names, dataset_names_temp)
        }
    }
}
```

## Perform motif search

```{r search}
# set up parallel processing
# create cluster
cores = 8
cl <- makeCluster(cores)
# get library support needed to run the code
clusterEvalQ(cl, {library(MItools); library(rtracklayer)})
# put objects in place that might be needed for the code
clusterExport(cl, c("dataset_names", datasets ,"myPPInetwork2SLIMFinder"), envir=environment())

R_data = parSapply(cl = cl, X = 1:nrow(dataset_names), FUN = function(i){
    myPPInetwork2SLIMFinder(dataset_name = dataset_names$dataset_names[i], slimlen = dataset_names$slimlen[i], maxwild = dataset_names$maxwild[i],
                            interaction_main_set = eval(parse(text = dataset_names$datasets[i])),
                            center_domains = dataset_names$domain_centered[i], analysis_type = dataset_names$analysis_type[i])
}, simplify = TRUE, USE.NAMES = TRUE)

stopCluster(cl)

# paths to RData files containing all objects created by all runs of this pipeline
print(R_data)
```