---
title: "Motif - Domain Results Table: motif - viral protein - human domain interactions"
author: "Vitalii Kleshchevnikov"
date: "17/08/2017"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
packages = c("MItools", "DT")
if(mean(packages %in% names(installed.packages()[,"Package"])) != 1){
    packages_to_install = packages[!packages %in% names(installed.packages()[,"Package"])]
    # specifying mirror is necessary for some Linux systems
    install.packages(packages_to_install, dependencies = T, repos = "http://mirrors.ebi.ac.uk/CRAN/")
    packages_to_install = packages[!packages %in% names(installed.packages()[,"Package"])]
    source("https://bioconductor.org/biocLite.R")
    biocLite(packages_to_install, dependencies = T)
    devtools::install_github("vitkl/MItools", dependencies = T)
}

suppressPackageStartupMessages({
    library(MItools)
    library(DT)
})
```


```{r load}
# Load results
name = "./DomainResultsTable_Vidal20171019.RData"
if(!file.exists(name)){
    load("./processed_data_files/what_we_find_VS_ELM_clust20171019.RData")
    occurence_QSLIMFinder = fread("./SLIMFinder_Vidal/result/occurence.txt")
    patterns_QSLIMFinder = fread("./SLIMFinder_Vidal/result/main_result.txt")
    comparimotif_wdb = fread("./SLIMFinder_Vidal/result/comparimotif.compare.tdt")
    rm(list = ls()[!ls() %in% c("XYZ.p.adjust","res_count", "sequential_filter", "name", "occurence_QSLIMFinder", "comparimotif_wdb", "patterns_QSLIMFinder")])
    save.image(file = name)
} else {
    load(name)
}
doman_viral_pairs = T # if false - add human proteins containing domains
motifs = T # based on Vidal's data
```

## How I identify Domains likely to mediate interaction?  

I estimate this using empirical p-value for seing a domain a number of times.  
More specifically, what is the chance of randomly seeing any domain the observed number of times among all proteins that interact with a specific viral protein?

```{r sigdomain_sigmotif}
fdr_motifs = 1
fdr_pval_thresh = 1
X = printMotifDomainTable(res_count, doman_viral_pairs = doman_viral_pairs, motifs = motifs, destfile = "./domains_fdr_corrected_only_with_significant_motifs_empirical_p_value.tsv",
               only_with_motifs = T, fdr_motifs = fdr_motifs, fdr_pval_thresh = fdr_pval_thresh,
               occurence_QSLIMFinder = occurence_QSLIMFinder,
               comparimotif_wdb = comparimotif_wdb,
               patterns_QSLIMFinder = patterns_QSLIMFinder, print_table = F)
```

## Counts of domains that are likely to mediate interaction with a viral protein that has an instance of a potentially new motif

More precisely, these domains are overrepresented in proteins that interact with a specific viral proteins. In addition to saying this domains may be mediating interaction, this can be also interpreted is a way that viral proteins interact with proteins that have functions conferred by the domain (if an enriched domain is an ATPase domain, then the viral protein prefferentially interacts with ATPase-s).

The number of domain instances is shown per each domain:  

```{r}
X4 = X[is.na(motif_mapping)]
X4[, .(NviralProt = data.table::uniqueN(viral_interactor)), by = domain_name][order(NviralProt, decreasing = T)]
```

## Motifs that don't match ELM

Each motif may be interacting with one of the many domains that were found enriched.

Motif_Dataset tells that sequences of interacting partners of ID-1 proteins were searched for motifs while the search space was restricted using ID-2 (example "interactors_of.ID-1.ID-2").  
The number of motif instances is shown per each motif:  

```{r}
X4[, .(NviralProt = data.table::uniqueN(viral_interactor)), by = Motif_Pattern][order(NviralProt, decreasing = T)]
DT::datatable(X4, escape = FALSE)
```